---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# categoryEncodings

<!-- badges: start -->
[![R-CMD-check](https://github.com/JSzitas/categoryEncodings/workflows/R-CMD-check/badge.svg)](https://github.com/JSzitas/categoryEncodings/actions)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/JSzitas/categoryEncodings?branch=master&svg=true)](https://ci.appveyor.com/project/JSzitas/categoryEncodings)
[![Codecov test coverage](https://codecov.io/gh/JSzitas/categoryEncodings/branch/master/graph/badge.svg)](https://codecov.io/gh/JSzitas/categoryEncodings?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![CRAN status](https://www.r-pkg.org/badges/version/categoryEncodings)](https://CRAN.R-project.org/package=categoryEncodings)
<!-- badges: end -->

**categoryEncodings** intends to provide a fast way to encode 'factor' or qualitative variables through various methods. The packages uses **data.table** as the backend for speed, with as few other dependencies as possible. Most of the methods are based on the paper of Johannemann et al.(2019) - Sufficient Representations for Categorical Variables (arXiv:1908.09874).

The current version features automatic inference of factors and uses a very simple heuristic for encoding, as well as allowing manual controls. 

## Installation
You can install the latest version of **categoryEncodings** from [github](https://github.com/JSzitas/categoryEncodings) using the *devtools* package

``` {r, eval = FALSE}
devtools::install_github("JSzitas/categoryEncodings")
```

**NOTE:** The latest stable version available from **CRAN** contains features deprecated in the current development version - I hope to resolve this soon, and publish the development version. 

## Example 

Here we want to encode all of the factors in a given data.frame.

````{r, message = FALSE, eval = TRUE}
library(categoryEncodings)
# create some example data
data_fm <- cbind( data.frame( 
   matrix( rnorm(5*100),ncol = 5)),
           sample(sample(letters, 10), 100, replace = TRUE),
           sample(sample(letters, 20), 100, replace = TRUE),
           sample(sample(1:10, 5), 100, replace = TRUE),
           sample(sample(1:50, 35), 100, replace = TRUE ),
           sample(1:2, 100, replace = TRUE ))
colnames(data_fm)[6:10] <- c( "few_letters",  "many_letters",
                              "some_numbers", "many_numbers",
                              "binary" ) 
# it does not matter how many factor variables there are, whether they are encoded as factors
# and whether you supply a method to encode them by - some simple inference of factors is done
# based on the number of distinct values in every variable - over a certain threshold 
# a variable is deemed as essentially a factor, and treated as such for conversion 
# you will be notified of which variables are being converted via a warning
result <- encoder(data_fm)
# note that due to the data.table back-end, the result has to be saved to an object to be 
# visible: otherwise printing is suppressed.   
head(result$encoded)
````

We also recover a **function closure** which we can reuse to fit new data, as long as it conforms to the same format: 
```{r}
# to fit to any dataset you can either call it directly - it is a single argument function
data_fm_encoded <- result$fitted_encoder(data_fm)

# or rename it, and stash it away for later use
encoding_function <- result$fitted_encoder
```

You also get a **"de-encoding"** function - 
```{r}
deencoder <- result$fitted_deencoder
```

This undoes the "encoding", effectively returning the original data. This can be quite useful for interpretability methods, where the 
interpretation becomes easier for un-encoded data. 
Note that this sadly does not maintain the order of the data from the original - and some attributes may be lost. Nonetheless, the recovered data is almost the same, and equivalent for all practical purposes: 

```{r}
original <- data.table::data.table(data_fm)
deencoded <- deencoder( result$encoded ) 


all.equal( data.table::setorder(original), 
           data.table::setorder(deencoded), 
           check.attributes = FALSE )
```

## Contributing

Please do contribute to the projects, all contributions are welcome, as long as people keep things civil - there is no need for negativity, hatred, and rudeness. Also, please do refrain from adding unnecessary dependencies (*Ex:* pipe) to the package (such pull requests as would add an unnecessary dependency will be denied/ suspended until the code can be made dependency free). This package wants to be as lightweight as possible - even if this means the code is a bit harder to write and maintain. 
